---
title: "Reproducible Research - Project 1"
author: "Timothy"
date: "May 19, 2016"
output: html_document
---

# Reproducible research peer assessment 1

Load in the data.

```{r echo=TRUE}
data <- read.csv("activity/activity.csv")
```

##What is the mean total number of steps taken per day?

We are to ignore missing values, i.e. NAs. Thus, where appropriate, na.rm will be set to **TRUE**


First, we use tapply() to obtaint the total number of steps for each day.
This new set of numbers can then by graphed in a histogram and we can use mean() and median() to get those respective statistics.
```{r, echo=TRUE}
stepstotal <- tapply(data$steps, data$date, sum, na.rm=TRUE)
hist(stepstotal,breaks=50,main = "Histogram of Steps")
```

```{r, echo=TRUE}
mean(stepstotal)
median(stepstotal)

```
The mean is `r mean(stepstotal)`
The median is `r median(stepstotal)`

##What is the average daily pattern?

```{r, echo=TRUE}
stepsinterval <- aggregate(steps ~ interval, data = data, FUN = mean, na.rm=TRUE)
head(stepsinterval)
plot(stepsinterval, type="l", main = "Average steps by interval", xlab ="5 minute interval", ylab = "Average steps across days")
```

```{r, echo=TRUE}
maxsteps <- max(stepsinterval$steps)
interval <- stepsinterval$interval[stepsinterval$steps==maxsteps]
```

The maximum number of steps is `r maxsteps` and occurss during the `r interval` interval.

###Imputing missing values

The number of NAs in the dataset:

```{r, echo=TRUE}
length(data$steps[data$steps == "NA"])
```

Fill in the NAs with the mean, and there is a handy function to impute called **impute()** in the package Hmisc
```{r, echo=TRUE}
library(Hmisc)
dataimpute <- data
dataimpute$steps <- impute(dataimpute$steps, fun=mean)
```
Make a histogram of total number of steps per day, report the mean and median.
```{r, echo=TRUE}
totalstepsimputed <- tapply(dataimpute$steps, dataimpute$date, sum)
hist(totalstepsimputed,breaks=50, main = "Histogram of Steps in imputed data")
```

```{r, echo=TRUE}
meanimputed <- mean(totalstepsimputed)
medianimputed <- median(totalstepsimputed)
```
The mean is `r meanimputed` and the median is `r medianimputed`
DO THEY DIFFER? WHAT IS THE IMPACT?

Yes, these differ from the originial values - mean and median are now the same, and have increased.

##Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
dataimpute$days <- weekdays(as.Date(dataimpute$date, format = "%Y-%m-%d"))

dataimpute$daytype <- ifelse(dataimpute$days== "Sunday" | dataimpute$days == "Saturday", "weekend", "weekday")
dataimpute$daytype <- as.factor(dataimpute$daytype)
```

```{r, echo=TRUE}
meanfinal <- aggregate(steps ~ interval + daytype, data = dataimpute, mean)
library(lattice)
xyplot(steps~interval|daytype, data=meanfinal, xlab="INTERVAL", ylab="NUMBER OF STEPS", type="l", layout=c(1,2))
```
